---
title: "Transform Data"
---

```{r setup}
library(tidyverse)
#install.packages("babynames")
library(babynames)
babynames
```

# dplyr

```{r}
ggplot(data = babynames) +
  geom_line(mapping = aes(x = year, y = prop))
```

## Your Turn 1

Alter the code to select just the `n` column:

```{r}
select(babynames, "n")
```

## Quiz

Which of these is NOT a way to select the `name` and `n` columns together?

```{r}
select(babynames, -c(year, sex, prop)) #deletes listed values, leaving n
select(babynames, name:n) #going to pull these 2 out becuase nothing else is between them
select(babynames, starts_with("n")) #names also starts with n, works because we want it
**select(babynames, ends_with("n"))**
```

## Your Turn 2

Use `filter`, `babynames`, and the logical operators to find:

* All of the names where prop is greater than or equal to 0.08  
* All of the children named "Sea"  

```{r}
filter(babynames, prop>=0.08)
filter(babynames, name == "Sea")
?babynames
```

## Your Turn 3

Use Boolean operators to return only the rows that contain:

* _Boys_ named Sue  
* Names that were used by exactly 5 or 6 children in 1880  
* Names that are one of Acura, Lexus, or Yugo

```{r}
filter(babynames, name == "Sue", sex == "M")
filter(babynames, (n == 5 | n == 6) & year == 1880)
filter(babynames, name =="Acura" | name =="Lexus" | name =="Yugo")
```

## Help Me

What is the smallest value of n?
What is the largest?

```{r}
arrange(babynames, n, prop) #prop is the tiebreaker (bigger pop in more recent years)
arrange (babynames, desc(n))
```

## Your Turn 4

Use `%>%` to write a sequence of functions that: 

1. Filters babynames to just the girls that were born in 2017, *then...*  
2. Selects the `name` and `n` columns, *then...*  
3. Arranges the results so that the most popular names are near the top.

```{r}
babynames %>%
  filter(year == 2017, sex == "F") %>%
  select(name, n) %>%
  arrange(desc(n))
```

## Your Turn 5 - Exam

1. Trim `babynames` to just the rows that contain **your** `name` and **your** `sex`  
2. Trim the result to just the columns that will appear in your graph (not strictly necessary, but useful practice)  
3. Plot the results as a line graph with `year` on the x axis and `prop` on the y axis

```{r}

```

## Your Turn 6

Copmplete the code below to extract the rows where `name == "Khaleesi"`. Then use `summarise()` and `sum()` and `min()` to find:

1. The total number of children named Khaleesi
2. The first year Khaleesi appeared in the data

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
babynames ___ 
  filter(_______________________) ___
  ___________(total = ________, first = _______)
```

## Your Turn 7

Use `group_by()`, `summarise()`, and `arrange()` to display the ten most popular names. Compute popularity as the *total* number of children of a single gender given a name.

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
babynames %>%
  _______(name, sex) %>% 
  _______(total = _____(n)) %>% 
  _______(desc(_____))
```


## Your Turn 8

Fill in the blanks to: 
Rank each name _within its year and sex_. 

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
babynames %>% 
  group_by(____, ____) %>% 
  ______(rank = min_rank(____(prop)))
```


***

# Take aways

* Extract variables with `select()`  
* Extract cases with `filter()`  
* Arrange cases, with `arrange()`  

* Make tables of summaries with `summarise()`  
* Make new variables, with `mutate()`  
* Do groupwise operations with `group_by()`

* Connect operations with `%>%`  


